### 2014-11-18 docs/ms.Rmd
Incoperated some of james comments. But ms has changed quite a bit from his version and so some were not relevant

Measuring reflectance spectra via spectrometry is now a reasonably inexpensive and convenient technique. Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008]. Experience, guidance and objectivity (e.g. double-blind experimental design) are necessary to ensure data are consistent and representative of the subject. Crucially, the conditions under which data were collected must be reported exactly to ensure repeatability.

Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration. High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013]. The main challenge for gathering colour data from photographs are in ensuring correct standardization - such as linearization - and consistent lighting [@stevens_using_2007]. Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013].

@@

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.

@@

We developed a list of metrics we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrophotometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2).

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 top colour science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Ecology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_). For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages. We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals. We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern; and 4) papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors. Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting. We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.

Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis. Overall, `r specs+both` studies used a spectrometer (henceforth "spec") to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spec and a camera. Reporting of the make and model of the spec and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was from the colour patch.

@@

Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used a hexagonal (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach. Case specific models were used in `r vismod_type[2]` papers. Of visual modelling papers, `r round(prop_irrad, digits =2)`% reported the type of irradiance when necessary, `r round(prop_spvismod, digits = 2)`% the species modelled as a receiver, and `r round(propvis_mod_param, digits = 2)`% made explicit and complete mention of the parameters they used (e.g. cone densities, signal-to-noise ratios). The background used in visual models was reported `r round(prop_bkg, digits = 2)`% of the time. The type of receptor noise modelled was reported in only `r round(prop_noise, digits = 2)`% of the studies. Explicit mention of the type of quantum catch (`r round(prop_qcatch, digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was modelled, was not common.

@@

Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration. While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details which are essential to the transparency and reproducibility of their work.  While all criteria are important, some deserve special attention. In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used. More importantly, the probe distance and angle should be standard reporting for all studies using spectrophometry. Visual models were regularly used to analyse spectral data, something we welcome. However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models. Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities. These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. It is said that a picture is worth 1000 words, and we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing. Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

Our objective in this paper was not to point out negative aspects of single studies, but to highlight an important gap in reporting that can influence the reproducibility of colour research and cross-study comparisons. By no means do we wish to say that studies lacking important details are not of significant value. Indeed, even in our own work [e.g. @umbers_bright_2013] we have failed to report important details, which in retrospect can lead to confusion about how such data were collected and analysed. This review of reporting is aimed at providing constructive criticism; it is only through critical appraisal and collegial discussion that we can identify such gaps as areas for improvement, and constructively contribute to the growth of our field. Such assessments are particularly important in an age where accountability, transparency and reproducibility are of utmost importance.

>>>>

Measuring reflectance spectra via spectrometry is now a reasonably inexpensive and convenient technique. Spectrometry allows the capture of quantitative and arguably repeatable data and has been rapidly taken up by researchers the field. However one drawback to this methodology, is the ease with which erroneous data can be collected if attention is not paid to finer details of spectrum capture. The angle from which spectra are taken [@santos_strong_2007], consistent calibration of light and dark standards, the distance between light source and surface, and the size, shape, glossiness or iridescence of the specimen can all contribute to variation in data and potentially bias results [@andersson_quantifying_2006; @kemp_ornamental_2008; @meadows_quantifying_2011; @schaefer_birds_2008]. Experience, guidance, consistency and objectivity (e.g. double-blind experimental design) are necessary to ensure data are repeatable and representative of the subject. Crucially, to ensure repeatability across studies, the conditions under which data were collected must be reported exactly.

Complementary to spectrometry, digital photography is increasingly being used to quantify biological colouration. High-resolution cameras are inexpensive,  and allow instantaneous, repeatable sampling of multiple colour patches [@mckay_use_2013]. The main challenge for gathering colour data from photographs involve ensuring standardization, such as linearization and consistent lighting [@stevens_using_2007]. Most digital cameras used for the measurement of biological colouration are consumer products designed for the human visual system. Thus hardware modifications for ultraviolet and infrared imaging are generally necessary [e.g. @stevens_colour_2013].

@@

Since the repeatability and transparency of science is paramount, it is imperative that we regularly and constructively reflect on our own practices [@alsheikh_public_2011; @sandve_ten_2013]. The potential for subjectivity in colour measurement and the mathematical opacity of visual models for the average biologist means that, as in any field, is it critical that methods are reported rigorously. Methods reporting should ideally have well defined standards such that sufficient detail is provided to allow comprehensive peer review, precise replication, cross-study comparison and data sharing [@alsheikh_public_2011]. The aims of this paper were to (a) review methods reporting in papers published on colour in 2013 from 22 top tier journals and (b) provide a set of recommendations for best-practice in reporting on the most common methods. Our review specifically focuses on papers that quantify the external coloration of a plant or animal. These may be purely descriptive quantifications of color, or they may be incorporated into models of animal vision. Our hope is that this paper will provide a useful set of guidelines to improve the reproducibility and transparency of studies on colour.

@@

We developed a list of metrics we deemed critical to be reported for reproducibility of colour methods (Table 1) and analysis (Table 2), loosely based on Andersson and Prager [-@andersson_quantifying_2006] and Montgomerie [@montgomerie_analyzing_2006]. These included quantitative methods including spectrophotometric and photographic techniques [as in @kemp_ornamental_2008; @chiao_visualization_2009] but not qualitative methods such as the use of Munsell chips or paint swatches [as in @hogstad_gloger_2009]. For visual modeling techniques we characterized a set of criteria required for reproducing the analysis (Table 2). Much of these criteria specifically targeted the necessary parameters required to be explicitly defined by the user during analysis. These are essential if such analyses are to be reproduced.

@@

To determine the current state of methods reporting in the field we searched for papers from 2013 in 22 top colour science journals (_Animal Behaviour_, _Behavioral Ecology_, _Journal of Experimental Biology_, _Biology Letters_, _Proceedings of the Royal Society B: Biological Sciences_, _Biological Journal of the Linnean Society_, _Ethology_, _The American Naturalist_, _Ecology Letters_, _Journal of Evolutionary Biology_, _American Journal of Botany_, _Ecology and Evolution_, _Oikos_, _Ecology_, _Journal of Ecology_, _Behavioral Ecology and Sociobiology_, _Current Zoology_, _Naturwissenschaften_, _New Phytologist_, _PLoS ONE_, _Evolution_, _Functional Ecology_). For each of the journals, we used the Boolean search terms 'colour*', 'color*' or 'spectra*' and searched for these in the title or abstract from each of the journal home pages. We read the titles and abstracts of each paper and included only those papers that explicitly used either a spectrophotometer or camera to quantify colour signals. We excluded papers that were: 1) review papers; 2) methodological papers; 3) papers quantifying colour pattern rather than characteristics of each individual colour patch; and 4) papers taking microspectrophotometric measurments of retinal absorbance. To review the large number of papers returned from the searches, journals were haphazardly divided up between the authors. In order to reduce the risk of observer bias in our scoring, papers were read and reassessed by a total of three authors. Any discrepancies between the assessor scores were discussed and agreed upon prior to analysis. We then tabulated what details of the methods were reported against our criteria. The data along with analysis script can be found at: https://github.com/daniel1noble/colsci_rep (TODO: GIVE IT A DOI?). We have kept the papers used in our dataset annoymous to avoid accusing specific colleagues of inadvertant reporting. We feel that the issue of reporting is something attributable to the ambiguity and lack of guidelines in our field and not something specific to individual research groups.

Our search returned a total of 216 papers, however, only `r nrow(coldat)` met our criteria and were included in the final analysis. Overall, `r specs+both` studies used a spectrometer (henceforth "spec") to measure colour, while `r cam+both` used a camera. Of these studies, `r both` quantified colour using both a spec and a camera. Reporting of the make and model of the spec and camera was high (Figure 1, spectrometers = `r round(specModel, digits = 2)`% and cameras = `r round(camModel, digits = 2)`%), however, only `r round(CamLight, digits = 2)`%  and `r round(specLight, digits = 2)`% of camera and spec studies reported a light source. In addition, the number of reflectance spectra and pixels averaged were only reported in `r round(specAVG, digits = 2)`% and `r round(camAVG, digits = 2)`% of studies, respectively. Surprisingly, dark standards were only reported in `r round(drk_std, digits = 2)`% of the spec studies, while white standards were reported in `r round(white_stdspec, digits = 2)`%. Studies quantifying spectral reflectance often did not report the integration time of the spectrophotmeter (`r round(intTime, digits = 2)`%), or the angle of the optical probe relative to the surface (`r round(angle, digits = 2)`%) or distance (`r round(dist, digits = 2)`%) the probe was held from the measurement surface.

@@

Most studies using visual models analysed their data using receptor-noise models (`r vismod_type[3]` of `r sum(vismod_type)` studies), while others used the colour hexagon (`r vismod_type[1]` studies), or tetrahedral colour space (`r vismod_type[4]` studies) approach. Case specific models were used in `r vismod_type[2]` papers. Of visual modelling papers, `r round(prop_irrad, digits =2)`% reported the type of irradiance when necessary, `r round(prop_spvismod, digits = 2)`% the species modelled as a receiver, and `r round(propvis_mod_param, digits = 2)`% made explicit and complete mention of the parameters they used (e.g. cone densities, signal-to-noise ratios). The background used in visual models was reported `r round(prop_bkg, digits = 2)`% of the time. The type of receptor noise modelled was reported in only `r round(prop_noise, digits = 2)`% of the studies. Explicit mention of the type of quantum catch (`r round(prop_qcatch, digits = 2)`% of studies), or whether photoreceptor adaptation (`r round(prop_adap, digits = 2)`%) was modelled, was not common.

@@

Our review of recent colour literature highlights some important reporting gaps that will hinder the reproducibility of studies of biological colouration. While we recognize that journals have space limitations and that work may reference previous methological and analytical details, we strongly believe that researchers can do a better job of reporting details that are essential to both the outcome of their results and reproducibility of their work.  While all criteria are important, some deserve special attention. In particular, we found that colour scientists could do a better job reporting light sources, the numbers of pixels or spectra averaged, and the type of standards used (i.e. white and black reference samples). More importantly, the distance and angle of the optical probe relative to the measurment surface should be standard reporting for all studies using spectrophometry. Visual models were regularly used to analyse spectral data, something we welcome. However, there were many important aspects of the analysis that were seldom reported, including the type of quantum catch used, whether photoreceptor adaptation was modelled and the type of neural noise used in receptor-noise models. Photoreceptor densities should also be reported where possible even if using standard avian photorecptor densities. These all have important consequences on downstream analysis and the reproducibility of colour signal quantification [e.g. @kemp_ornamental_2008; @santos_comparison_2007]. In addition, we recommend the presentation of spectral reflectance curves where possible; these figures are visually appealing and allow for the rapid assesment of the nature and quality of the data, independent of downstream processing. Surprisingly, we found that only `r round(fig, digits = 2)`% of studies using spectral reflectance data had a figure showing these curves. When possible, we also recommend that authors justify data capture and analytical approaches - something few studies we investigated explicitly outlined.

Our objective in this paper was not to highlight shortcomings in single studies, but to bring attantion to common oversights in reporting that can influence the reproducibility of colour research and cross-study comparisons. By no means do we wish to say that studies lacking important details are not of significant value. Indeed, even in our own work [e.g. @umbers_bright_2013] we have failed to report important details, which in retrospect can lead to confusion about how such data were collected and analysed. This review of reporting is aimed at providing constructive criticism; it is only through critical appraisal and collegial discussion that we can identify such gaps as areas for improvement, and constructively contribute to the growth of our field. Such assessments are particularly important in an age where accountability, transparency and reproducibility are of utmost importance.
